Simple repression draft

### 2020-05-30 doc/appendix_03_bayesian_inference.tex
MRM progress towards cleaning the simple-repression inference section in the SI (with new figure)

\mrm{Still need to write something about the intuition of being data-rich and 
how then the prior matters little.}

@@

With this, we are ready to write our full generative model as
\begin{equation}
\begin{split}
\ln k_i \sim \text{Normal}(-0.5, 2),
\\
\ln b \sim \text{Normal}(0.5, 1),
\\
m \sim \text{NBinom}(k_i, b).
\end{split}
\end{equation}
Section ~\ref{section_04_bayesian_inference} in the main text details the
results of applying this inference to the single-cell mRNA counts data.

@@

Now that we have a satisfactory model in hand for constitutive
promoters, we would like to return to the main thread: can we
infer repressor binding and unbinding rates from mRNA distributions
over a population of cells as measured by FISH in~\cite{Jones2014}?
If so, how do these inferred rates compare to direct
single-molecule measurements such as from~\cite{Hammar2014} and
to binding energies such as from~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018}, which were inferred under the assumptions
of the equilibrium models in Figure~\ref{fig1:means_cartoons}(B)?
And can this comparison shed light on the unreasonable effectiveness
of the equilibrium models?

>>>>

For our particular case all of the datasets from~\cite{Jones2014} used in this
paper have $\mathcal{O}(10^3)$ data points. What this implies is that our
particular choice of priors will not significantly affect our inference as long
as they are broad enough. A way to see why this is the case is to simply look at
Bayes theorem. For $N ~ 1000-3000$ datum all of the independent of each other
and $n \ll 10^3$ parameters Bayes theorem reads as
\begin{equation}
P(\vec{\theta} \mid D) \propto \prod_{k=1}^{N} P(d_k \mid \vec{\theta})
\prod_{i=1}^n P(\theta_i),
\end{equation}
where $d_k$ represents the $k$-th datum. That means that if our priors span a
wide range of parameter space, the posterior distribution would be dominated by
the likelihood function.

@@

With this, we are ready to write our full generative model as
\begin{equation}
\begin{split}
\ln k_i \sim \text{Normal}(-0.5, 2),
\\
\ln b \sim \text{Normal}(0.5, 1),
\\
m \sim \text{NBinom}(k_i, b).
\end{split}
\end{equation}
Section~\ref{section_04_bayesian_inference} in the main text details the
results of applying this inference to the single-cell mRNA counts data.

@@

As detailed in~\ref{section_04_bayesian_inference} in the main text the
inference on the unregulated promoter served as a stepping stone towards our
ultimate goal of inferring repressor rates from the steady-state mRNA
distributions of simple-repression architectures. For this we expand the
one-state bursty promoter model to a two-state promoter as schematized in
Figure~\ref{fig1:means_cartoons}(C) as model 5. This model adds two new
parameters: the repressor binding rate $k^+$, solely function of the repressor
concentration, and the repressor dissociation rate $k^-$, solely a function of
the repressor-DNA binding affinity.

The structure of the data in~\cite{Jones2014} for regulated promoters tuned
these two parameters independently. In their work the production of the LacI
repressor was under the control of an inducible promoter regulated by the TetR
repressor as schematized in Figre~\ref{figS:aTc_circuit}. When TetR binds to the
small molecule anhydrotetracycline (aTc), it shifts to an inactive conformation
unable to bind to the DNA. This translates into an increase in gene expression
level. In other words, the higher the concentration of aTc added to the media,
the less TetR repressors that can control the expression of the \textit{lacI}
gene, so the higher the concentration of LacI repressors in the cell.

\begin{figure}[h!]
\centering
\includegraphics{../figures/si/figS0X_aTc_circuit.pdf}
\caption{\textbf{aTc controlled expression of LacI repressor.} Schematic of the
circuit used in~\cite{Jones2014} to control the expression of the LacI
repressor. The \textit{lacI} gene is under the control of the TetR repressor. As
the TetR repressor is inactivated upon binding of anhydrotetracycline or aTc,
the more aTc added to the media were cells are growing, the less TetR repressors
available to control the expression of the \textit{lacI} gene, resulting in more
LacI repressors per cell. LacI simultaneously controls the expression of the
mRNA on which single-molecule mRNA FISH was performed for gene expression
quantification.}
\label{figS:aTc_circuit}
\end{figure}
